Basic Scatter Plot (*Nothing Very Fancy - yet*)
===============================================

We'll be using the `diamonds` dataset that comes with `ggplot2`.

###Loading Required Packages###
```{r,'Loading Required Packages',R.options=list(width=160),echo=FALSE}

require(ggplot2)
require(gridExtra)
```

###Basic Data Summary###

```{r,'Basic Data Summary',R.options=list(width=160)}

diamonds<-diamonds
summary(diamonds)
colnames(diamonds)
set.seed(56)
data=diamonds[sample(1:nrow(diamonds),2000),]
```


###Plots###

The `ggplot2` package by *Hadley Wickham* follows a layer by layer plot building philosophy. In order to build a plot using `ggplot2` , the ideal way is to think in terms of plotting layers.

**The steps involved in creating a plot are as follows:**

    +**Step 1:** Add dataset:         `ggplot(data)`
    +**Step 2:** Add geometry(layer): `geom_point(x,y)`
    +**Step 3:** Add Plot title/change axes scales/faceting

####Plot 01####


**Create a basic scatter plot:**

Here:

`data` is a sample of 2000 observations from diamonds dataset.

The `null` plot is initialised by using `ggplot()`

To create a scatterplot, we use `geom_point()` as geometry/layer

To add plot title we use `ggtitle()`

```{r,fig.width=13,fig.height=13,'Basic Scatter Plot'}

p1=ggplot(data)+ 
  geom_point(aes(y=price,x=carat))+
  ggtitle('p1: Basic scatter plot on diamonds dataset - Price vs Carats')

p1

```

####Plot 02####

To create a scatter plot with colors according to a given factor variable, we use the `color` parameter. It takes in a factor (categorical) variable as input & colors the points according to the components of factor variable.

```{r,fig.width=13,fig.height=13,'Scatter Plot-Color added'}

p2=ggplot(data)+
  geom_point(aes(y=price,x=carat,color=cut))+
  ggtitle('p1: Scatter plot on diamonds dataset with colors by "cut"- Price vs Carats')

p2

```

The above plot is exactly same as `p2=p1+geom_point(aes(x=carat,y=price,color=cut))`. If you already have a plot & intend to modify its one aspect using `ggplot2`, you can do it by using the `+` operator along with your plot's layer/geometry specification.


####Plot 03####

```{r,fig.width=13,fig.height=13,'Scatter Plot-Shape added'}
p3=ggplot(data)+
  geom_point(aes(y=price,x=carat,color=cut,shape=cut))+
  ggtitle('p3: Scatter plot with colour & Shape parameters')

p3

```

####Plot 04####

```{r,fig.width=13,fig.height=13,'Scatter Plot-Size added'}

p4=ggplot(data)+
  geom_point(aes(y=price,x=carat,color=cut,size=depth))+
  ggtitle('p4: Scatter plot with colour & Size (by depth) parameters')

p4
```

####Plot 05: Combined plots####

```{r,fig.width=19,fig.height=20,'Combined Plots'}

grid.arrange(p1,p2,p3,p4,ncol=2,main='Scatter Plots')

```


####Plot 06: Adding Faceting grid####

```{r,fig.width=13,fig.height=13,'Adding a variety of paramters',fig.width=20,fig.height=20}

ggplot(data)+
  geom_point(aes(x=carat,y=price,color=cut),alpha=0.7)+
  facet_grid(facets=cut~clarity)+
  ggtitle('More parameters added to the scatterplot')

```